Progress Report on the ``Adult Database'' Analysis

### 2009-11-12 report/paper.tex
added references and citations.

Classification in data mining is the activity of identifying a given object as 
belonging to a pre-defined class. The ultimate goal is to learn a concept. This
learning process is based on past observations, where the dataset performs a
main role. This report gives a first overview on the ``Adult Database'' dataset,
describing its format and identifying the problem it poses. Some data analysis
is then done, in order to obtain some characteristics on the global set of
values. Finally, the future work plan is presented and some conclusions on the
subject are taken.

@@

One important objective in classifying things is to understand why the given object
is put under one class and not in another, that is, we want to understand what social characteristics
allows one person to gain more than someone else, if the sex is important, the education level,
the race and what factors are less important.

@@

Another method to analyze the dataset is based on the generation of clusters. 
Clusters are sets of data in which objects in the same cluster are similar, and
objects in different clusters are dissimilar. The first attempt at clustering
used only three variables: age, education\_num and hours\_per\_week, the same
variables we saw that had a greater impact on the amount gained by the
individual. The data was normalized using the Z-Transformation method and the 
K-Means method was used to obtain the final clusters, with a k of 6. Figures 
\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
scatter plots of the variables, colored by clusters. One thing one is able to
note is that people with younger age belong to cluster \#2. In terms of
education, people with higher education belong to cluster \#0, while people with
lower education belong to cluster \#1. In terms of hours of work per week, 
distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
shows how clusters relate to the class values. Although the distribution is 
balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
50000 dollars. Notice, as stated before, that this were the clusters having 
people with lower education and younger age, respectively.

@@

\begin{thebibliography}{1}
\end{thebibliography}

>>>>

Classification in data mining is the activity of identifying a given object as 
belonging to a pre-defined class \cite{1}. The ultimate goal is to learn a concept. This
learning process is based on past observations, where the dataset performs a
main role. This report gives a first overview on the ``Adult Database'' dataset,
describing its format and identifying the problem it poses. Some data analysis
is then done, in order to obtain some characteristics on the global set of
values. Finally, the future work plan is presented and some conclusions on the
subject are taken.

@@

One important objective in classifying things is to understand why the given object
is put under one class and not in another \cite{2}\cite{3}, that is, we want to understand what social characteristics
allows one person to gain more than someone else, if the sex is important, the education level,
the race and what factors are less important.

@@

Another method to analyze the dataset is based on the generation of clusters. 
Clusters are sets of data in which objects in the same cluster are similar, and
objects in different clusters are dissimilar. The first attempt at clustering
used only three variables: age, education\_num and hours\_per\_week, the same
variables we saw that had a greater impact on the amount gained by the
individual. The data was normalized using the Z-Transformation method and the 
K-Means method was used to obtain the final clusters, with a k of 6, using RapidMiner \cite{4}. Figures 
\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
scatter plots of the variables, colored by clusters. One thing one is able to
note is that people with younger age belong to cluster \#2. In terms of
education, people with higher education belong to cluster \#0, while people with
lower education belong to cluster \#1. In terms of hours of work per week, 
distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
shows how clusters relate to the class values. Although the distribution is 
balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
50000 dollars. Notice, as stated before, that this were the clusters having 
people with lower education and younger age, respectively.

@@

\begin{thebibliography}{1}
\bibitem{1}
Mitchel, T.; Machine Learning; McGraw-Hill, 1997
\bibitem{2}
Han, J., Kamber, M., Pei, J.; Data Mining: Concepts and Techniques, Second Edition; Morgan Kaufmann, 2005
\bibitem{3}
H. Witten, I., Frank, E.; Data Mining: Practical Machine Learning Tools and Techniques, Second Edition; Morgan Kaufmann, 2005
\bibitem{4}
RapidMiner; \url{http://www.rapidminer.com/}, November 2009
\bibitem{5}
Ross Quinlan, J.; C4.5: Programs for Machine Learning; Morgan Kaufmann, 1992
\end{thebibliography}
